### 2023-05-05 Diagnosis_Vis.Rmd

---
title: "Diagnosis Visualizations"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning = FALSE)
# Libraries 
library(tidyverse)
library(dplyr)
library(haven)
library(ggplot2)
library(tidyr)
library("lubridate")

#Data
Mothership<-read.csv("~/Desktop/Coding/data/Mothership_Vis.csv")
Mothership<-mutate(Mothership,
      DC_status_1 = as.factor(DC_status_1),
      Education_1 = as.factor(Education_1),
      Gender_1 = as.factor(Gender_1),
      Relationship_1 = as.factor(Relationship_1),
      Sexuality_1 = as.factor(Sexuality_1),
      IPT_track_1 = as.factor(IPT_track_1),
      Race_1 = as.factor(Race_1),
      Education_1 = as.factor(Education_1),
      Duration = as.factor(Duration),
      )

Mothership$Intake_1<-ymd(Mothership$Intake_1,locale = "en_US.UTF-8")
Mothership$TxYear<-as.numeric(format(Mothership$Intake_1,"%Y"))
Mothership<-mutate(Mothership,
                   TxYear = as.factor(TxYear),
                   TxYear = recode(TxYear,
                                   "1582"="2014",
                                   "2021"="2020",
                                   "2022"="2020"))
```

The following code creates visualizations for diagnostic related information. The main difficulty of these visualizations was getting diagnostic variables correctly coded so there was not overlap or double acounting for conditions.

# Create Diagnosis Variables 
```{r}
Mothership<-Mothership %>%
  mutate(
          # Mood Disorders
                     Mood_D = as.factor(
                      ifelse(
                        MDD_P ==1 | Bipolar1_P == 1 | Bipolar2_P == 1 | Dysthymia_P == 1, 1, 0)),
          # Anxiety Disorders
                    Anxiety_D = as.factor(
                      ifelse(
                        GAD_P ==1 | Panic_P == 1 | PanicAgor_P == 1 | Agoraphobia_P == 1 |
                        Social_P == 1 | Specific_P == 1, 1, 0)),
          # Obsessive Compusive Disorders
                #OCD_P is for current OCD
          # Personality Disorders
                #BPD_P is for current BPD
          # Substance Use Disorders
                    Substance_D = as.factor(
                      ifelse(
                        Alcohol_P == 1 | Drug_P ==1, 1, 0)),
          # Trauma or stresor related disorders
                    Trauma_D = as.factor(
                      ifelse(PTSD_P == 1 | Adjustment_P ==1, 1,0)),
          # Psychotic Disorders
                    Psychotic_D = as.factor(
                      ifelse(
                        Schiz_P == 1 | Schizaff_P == 1,1,0)),
            # Eating Disorders
                    Eating_D = as.factor(
                      ifelse(
                        Anorex_P == 1 | Bulim_P == 1 | Binge_P == 1, 1,0)),
            # Somatic Disroders
                    Somatic_D = as.factor(
                      ifelse(
                        Somat_P == 1 | UndiffSoma_P == 1| Hypochondriasis_P ==1,
                        1,0)),
                    Disorder_Type = as.factor(
                      ifelse(Anxiety_D == 1, "Anxiety",
                      ifelse(OCD_P == 1, "OCD",
                      ifelse(BPD_P == 1, "BPD",
                      ifelse(Substance_D == 1, "Substance Use",
                      ifelse(Trauma_D == 1, "Trauma",
                      ifelse(Mood_D == 1, "Mood",
                      ifelse(Psychotic_D == 1, "Psychotic", 
                      ifelse(Eating_D == 1, "Eating", 
                      ifelse(Somatic_D == 1, "Somatic", NA
                      )))))))))),
                  Anxiety_C =  as.factor(
                    ifelse( GAD_P !=1 & GAD_C ==1 | 
                        Panic_P != 1 & Panic_C == 1 | 
                        PanicAgor_P != 1 & PanicAgor_C == 1 | 
                        Agoraphobia_P != 1 & Agoraphobia_C == 1 |
                        Social_P != 1 & Social_C == 1 | 
                        Specific_P != 1 & Specific_C == 1, 1, 0)),
                  Mood_C =  as.factor(
                    ifelse( MDD_P != 1 & MDD_C == 1 |
                        Bipolar1_P != 1 & Bipolar1_C == 1 | 
                        Bipolar2_P != 1 &  Bipolar2_C == 1 | 
                        Dysthymia_P != 1 & Dysthymia_C == 1 , 1, 0)),
                  Substance_C = as.factor(
                    ifelse(Alcohol_P != 1 & Alcohol_C == 1 | 
                        Drug_P != 1 & Drug_C ==1, 1, 0)),
                  Trauma_C = as.factor(
                    ifelse(PTSD_P != 1 & PTSD_C == 1 | 
                        Adjustment_P != 1 & Adjustment_C ==1, 1,0)),
                  Psychotic_C = as.factor(
                    ifelse(Schiz_P != 1 & Schiz_C == 1 | 
                        Schizaff_P != 1 & Schizaff_C == 1,1,0)),
                  Eating_C = as.factor(
                    ifelse(Anorex_P != 1 & Anorex_C == 1 | 
                        Bulim_P != 1 & Bulim_C == 1 | 
                        Binge_P != 1 & Binge_C == 1, 1,0)),
                  Somatic_C = as.factor(
                    ifelse(Somat_P != 1 & Somat_C == 1 | 
                        UndiffSoma_P != 1 & UndiffSoma_C == 1 | 
                        Hypochondriasis_P != 1 &  Hypochondriasis_C == 1, 1,0)),
                  OCD_Co_C = as.factor(
                    ifelse(OCD_P != 1 & OCD_C == 1,1,0)),
                  BPD_Co_C = as.factor(
                    ifelse(BPD_P != 1 & BPD_C ==1,1,0)),
          #emotional disorders
                  Emotional_dis = as.factor(
                    ifelse(Anxiety_D == 1 | OCD_P == 1 | Trauma_D == 1 |
                           Mood_D == 1, "Emotional Disorder", "Other Disorder")),
                  Co_Anx = as.factor(
                        ifelse(Anxiety_D == 1 & Mood_C == 1 |
                                Anxiety_D == 1 & OCD_C == 1 |
                                Anxiety_D == 1 & BPD_C == 1 |
                                Anxiety_D == 1 & Substance_C == 1 |
                                Anxiety_D == 1 & Trauma_C == 1 |
                                Anxiety_D == 1 & Eating_C ==1 |
                                Anxiety_D == 1 & Psychotic_C ==1 |
                                Anxiety_D == 1 & Somatic_C ==1, 1, 0)),
                  Co_Mood = as.factor(
                        ifelse(Mood_D == 1 & Anxiety_C == 1 |
                                Mood_D == 1 & OCD_C == 1 |
                                Mood_D == 1 & BPD_C == 1 |
                                Mood_D == 1 & Substance_C == 1 |
                                Mood_D == 1 & Trauma_C == 1 |
                                Mood_D == 1 & Eating_C ==1 |
                                Mood_D == 1 & Psychotic_C ==1 |
                                Mood_D == 1 & Somatic_C ==1 , 1, 0)),
                  Co_OCD = as.factor(
                        ifelse(OCD_C == 1 & Anxiety_C == 1 |
                                OCD_C == 1 & Mood_C == 1 |
                                OCD_C == 1 & BPD_C == 1 |
                                OCD_C == 1 & Substance_C == 1 |
                                OCD_C == 1 & Trauma_C == 1 |
                                OCD_C == 1 & Eating_C ==1 |
                                OCD_C == 1 & Psychotic_C ==1 |
                                OCD_C == 1 & Somatic_C ==1, 1, 0)),
                  CO_BPD = as.factor(
                        ifelse(BPD_C == 1 & Anxiety_C == 1 |
                                BPD_C == 1 & Mood_C == 1 |
                                BPD_C == 1 & OCD_C == 1 |
                                BPD_C == 1 & Substance_C == 1 |
                                BPD_C == 1 & Trauma_C == 1 |
                                BPD_C == 1 & Eating_C ==1 |
                                BPD_C == 1 & Psychotic_C ==1 |
                                BPD_C == 1 & Somatic_C ==1 , 1, 0)),
                  CO_Psychotic = as.factor(
                       ifelse(Psychotic_D == 1 & Mood_C == 1 |
                                Psychotic_D == 1 & OCD_C == 1 |
                                Psychotic_D == 1 & BPD_C == 1 |
                                Psychotic_D == 1 & Substance_C == 1 |
                                Psychotic_D == 1 & Trauma_C == 1 |
                                Psychotic_D == 1 & Eating_C ==1 |
                                Psychotic_D == 1 & Anxiety_C ==1 |
                                Psychotic_D == 1 & Somatic_C ==1, 1, 0)),
                  CO_Eating = as.factor(
                      ifelse(Eating_D == 1 & Mood_C == 1 |
                                Eating_D == 1 & OCD_C == 1 |
                                Eating_D == 1 & BPD_C == 1 |
                                Eating_D == 1 & Substance_C == 1 |
                                Eating_D == 1 & Trauma_C == 1 |
                                Eating_D == 1 & Psychotic_C ==1 |
                                Eating_D == 1 & Anxiety_C ==1 |
                                Eating_D == 1 & Somatic_C ==1, 1, 0)),
                  CO_Somatic =  as.factor(
                       ifelse(Somatic_D == 1 & Mood_C == 1 |
                                Somatic_D == 1 & OCD_C == 1 |
                                Somatic_D == 1 & BPD_C == 1 |
                                Somatic_D == 1 & Substance_C == 1 |
                                Somatic_D == 1 & Trauma_C == 1 |
                                Somatic_D == 1 & Eating_C ==1 |
                                Somatic_D == 1 & Anxiety_C ==1 |
                                Somatic_D == 1 & Psychotic_C ==1, 1, 0)))

##### df for plots

# df without NAs in date and disorder type
Mothership_Diag <- Mothership %>%
  filter(!is.na(Disorder_Type))%>%
  filter(!is.na(TxYear))

# df with and without depression
No_depressy <- Mothership_Diag %>%
  filter(Disorder_Type != "Mood")
```

# Number of Disorders Treated Over Time

## {.tabset}

### All Disorders
```{r}
# Libraries
library(ggplot2)
#install.packages("hrbrthemes")
library(hrbrthemes)
library(dplyr)
library(tidyr)
library(viridis)

# Stacked
ggplot(data=subset(Mothership_Diag, !is.na(Disorder_Type)), 
       mapping= aes(x=TxYear, y = stat(count),
                    group=Disorder_Type, fill=Disorder_Type)) +
    geom_density(adjust=1.5,alpha=.4,position="fill") +
    theme(axis.text.x = element_text(angle = 45,hjust = 1))+
     theme(panel.grid.major.y = element_line(color = "grey",size = 0.5,linetype = 2))+
     xlab("Treatment Year") +
  theme(panel.background=NULL) +
    facet_wrap(~Emotional_dis)+
  labs(
    title = "Type of Disorders Treated at the Hospital Per Year",
    subtitle = "Broken Up By Disorder Types") +
    xlab("Treatment Year") +
    ylab("Percent of Patients With Disorder")

# Broken up by disorder. 
ggplot(data=subset(Mothership_Diag, !is.na(Disorder_Type)), 
    mapping= aes(x=TxYear,y = stat(count),
                 group=Disorder_Type, fill=Disorder_Type)) +
  geom_density(adjust=1.5,alpha=.6) +
  facet_wrap(~Disorder_Type)+
  theme_classic()+
    theme(
      panel.grid.major.y = element_line(color = "grey",size = 0.5,linetype = 2),
      axis.text.x = element_text(angle = 45,hjust = 1),
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      axis.ticks.x=element_blank(),
      panel.background=NULL) +
  labs(
    title = "Type of Disorders Treated at the Hospital Per Year",
    subtitle = "Broken Up By Dmotional vs Other Disorders") +
    xlab("Treatment Year") +
    ylab("Number of Patients With Disorder")

```
### Excluding Mood Disorders

```{r}
# without depression
ggplot(data=subset(No_depressy, !is.na(Disorder_Type)), 
    mapping= aes(x=TxYear,y = stat(count),
                 group=Disorder_Type, fill=Disorder_Type)) +
  geom_density(adjust=1.5,alpha=.6) +
  facet_wrap(~Disorder_Type)+
  theme_classic()+
    theme(
      panel.grid.major.y = element_line(color = "grey",size = 0.5,linetype = 2),
      axis.text.x = element_text(angle = 45,hjust = 1),
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      axis.ticks.x=element_blank(),
      panel.background=NULL) +
  labs(
    title = "Type of Disorders Treated at the Hospital Per Year",
    subtitle = "Broken Up By Dmotional vs Other Disorders") +
    xlab("Treatment Year") +
    ylab("Number of Patients With Disorder")
```

# Amount of Disorder Type Treated 
Grouped into DSM chapter subtypes (e.g., Mood Disorder = Major Depressive Disorder, Bipolar 1 & 2, etc.)

### Amount of Primary Diagnoses
```{r}
# Disorder Type 
ggplot(data = subset(Mothership_Diag, !is.na(Disorder_Type)),mapping = aes(x = Disorder_Type, fill=Disorder_Type))+
  geom_bar()+
  ggtitle("Frequency of Psychiatric Disorder Subtypes") +
  theme(axis.text.x = element_text(angle = 45,hjust = 1),
        legend.position = "none")+
  xlab("Psychiatric Disorder Subtype")+
  theme(panel.grid.major.x = element_line(color = "grey",size = 0.5,linetype = 2))+
  theme(panel.background=NULL)+
  scale_y_continuous(name = "Frequency",limits=c(0,3500))+
  coord_flip()
```

As you can see, the main disorder type in the hospital is mood realted disorders. This shouldn't be supprising as many indiviudals with suicidality (A common hospitalization issue) will meet criteria for a mood disorder.

### Excluding Mood Disorders
```{r}
# Disorder Type (Without Depression)
ggplot(data = subset(No_depressy, !is.na(Disorder_Type)),mapping = aes(x = Disorder_Type, fill=Disorder_Type))+
  geom_bar()+
  ggtitle("Frequency of Psychiatric Disorder Subtypes") +
  theme(axis.text.x = element_text(angle = 45,hjust = 1),
        legend.position = "none")+
  xlab("Psychiatric Disorder Subtype")+
  theme(panel.grid.major.x = element_line(color = "grey",size = 0.5,linetype = 2))+
  theme(panel.background=NULL)+
  scale_y_continuous(name = "Frequency",limits=c(0,3500))+
  coord_flip()
```

### Number of Co-occuring Disorders
The following visualization looks at the number of disorders people had, but wasn't the main reason they attended the hospital, or their "main" diagnosis.
```{r}
Mothership_Current_Long <- Mothership %>%
  pivot_longer(
    cols=c(Anxiety_C, Mood_C, Substance_C, Trauma_C, Psychotic_C, Eating_C,  Somatic_C, OCD_Co_C, BPD_Co_C),
    names_to = 'Disorder_Present') %>%
  filter(value != 0) %>%
  select(!value)  %>%
  filter(!is.na(Disorder_Present))%>%
  filter(!is.na(Disorder_Type))

ggplot(data = subset(Mothership_Current_Long, !is.na(Disorder_Present)),mapping = aes(x = Disorder_Present, fill=Disorder_Present))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 45,hjust = 1),
        legend.position = "none") +
  labs(title = "Number of Co Occuring Psychiatric Diosrders",
    subtitle = "In Addition to the Primary Diagnosis",
    x = "Co-Occuring Psychiatric Disorder") +
    scale_x_discrete(labels=c("Trauma_C" = "Trauma", 
                            "Substance_C" = "Substance Use",
                            "Somatic_C" = "Somatic",
                            "Psychotic_C" = "Psychotic",
                            "OCD_Co_C" = "OCD",
                            "Mood_C" = "Mood",
                            "Eating_C" = "Eating",
                            "BPD_Co_C" = "BPD",
                            "Anxiety_C" = "Anxiety")) +
  theme(panel.grid.major.x = element_line(color = "grey",size = 0.5,linetype = 2))+
  theme(panel.background=NULL)+
  scale_y_continuous(name = "Frequency",limits=c(0,3500))+
  coord_flip()
```

# Primary and Co-occuring Disorders.

The following visualization shows which disorders are co-occurring with which types of primary diagnoses. 
```{r}
library(dplyr)
library(tidyr)

Mothership_Current_Long %>%
  ggplot(aes(x = Disorder_Type, y = Disorder_Present, color = Disorder_Type)) +
  geom_count(alpha = .5)+
  scale_size_area(max_size = 20) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45,hjust = 1),
    legend.position = "none") +
  scale_y_discrete(labels=c("Trauma_C" = "Co Trauma", 
                            "Substance_C" = "Co Substance Use",
                            "Somatic_C" = "Co Somatic",
                            "Psychotic_C" = "Co Psychotic",
                            "OCD_Co_C" = "Co OCD",
                            "Mood_C" = "Co Mood",
                            "Eating_C" = "Co Eating",
                            "BPD_Co_C" = "Co BPD",
                            "Anxiety_C" = "Co Anxiety")) +
  labs(title = "Number of Co Occuring Psychiatric Diosrders per Primary Diagnosis",
    subtitle = "Grouped into subtypes of DSM diagonsis",
    y = "Co-Occuring Psychiatric Disorder",
    x = "Primary Psychiatric Diagnosis") 
```