### 2023-05-05 Treatment_Outcomes.Rmd

---
title: "Treatment Outcome Visualization"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning = FALSE)
# Libraries 
library(tidyverse)
library(dplyr)
library(haven)
library(ggplot2)
library(tidyr)
library("lubridate")

#Data
Mothership<-read.csv("~/Desktop/Coding/data/Mothership_Vis.csv")
Mothership<-mutate(Mothership,
      DC_status_1 = as.factor(DC_status_1),
      Education_1 = as.factor(Education_1),
      Gender_1 = as.factor(Gender_1),
      Relationship_1 = as.factor(Relationship_1),
      Sexuality_1 = as.factor(Sexuality_1),
      IPT_track_1 = as.factor(IPT_track_1),
      Race_1 = as.factor(Race_1),
      Education_1 = as.factor(Education_1),
      Duration = as.factor(Duration),
      )

Mothership$Intake_1<-ymd(Mothership$Intake_1,locale = "en_US.UTF-8")
Mothership$TxYear<-as.numeric(format(Mothership$Intake_1,"%Y"))
Mothership<-mutate(Mothership,
                   TxYear = as.factor(TxYear),
                   TxYear = recode(TxYear,
                                   "1582"="2014",
                                   "2021"="2020",
                                   "2022"="2020"))
```

# Remission from Depression Questionnaire (RDQ)

The following code visualizes the outcomes from the RDQ. The RDQ is broken up into the following categories:

1. Coping Skills: (higher score = more coping skills)
2. Positive Mental Health: (higher score = more positive mental health)
3. Functioning: (higher score = better functioning)
4. Well Being: (higher score = better overall well being)
5. Symptoms: (higher score = more severe depression symptoms)

```{r}
#Read in a New Data Set
Mothership_RDQ<- read_sav("~/Desktop/Coding/data/Mothership_DV.sav")
Mothership_RDQ<-select(Mothership,ID1:Sexuality_1,MDD_C:Day40_CUXOS)
Mothership_RDQ<-data.frame(sapply(Mothership_RDQ,FUN=as.numeric))

#Make This Data Set Long Data
library(tidyr)
Mothership_Long_RDQ <- gather(Mothership_RDQ,rm,RDQ,rdqPRE_coping:rdqPOST_sym,
                              factor_key = "T")
Mothership_Long_RDQ <- Mothership_Long_RDQ[order(Mothership_Long_RDQ$ID1),]

#Create a Time Variable
Mothership_Long_RDQ$PrePost <- NA
Mothership_Long_RDQ$PrePost[Mothership_Long_RDQ$rm=="rdqPRE_coping"]<-"Pre"
Mothership_Long_RDQ$PrePost[Mothership_Long_RDQ$rm=="rdqPRE_pmh"]<-"Pre"
Mothership_Long_RDQ$PrePost[Mothership_Long_RDQ$rm=="rdqPRE_func"]<-"Pre"
Mothership_Long_RDQ$PrePost[Mothership_Long_RDQ$rm=="rdqPRE_wbs"]<-"Pre"
Mothership_Long_RDQ$PrePost[Mothership_Long_RDQ$rm=="rdqPRE_sym"]<-"Pre"
Mothership_Long_RDQ$PrePost[Mothership_Long_RDQ$rm=="rdqPOST_coping"]<-"Post"
Mothership_Long_RDQ$PrePost[Mothership_Long_RDQ$rm=="rdqPOST_pmh"]<-"Post"
Mothership_Long_RDQ$PrePost[Mothership_Long_RDQ$rm=="rdqPOST_func"]<-"Post"
Mothership_Long_RDQ$PrePost[Mothership_Long_RDQ$rm=="rdqPOST_wbs"]<-"Post"
Mothership_Long_RDQ$PrePost[Mothership_Long_RDQ$rm=="rdqPOST_sym"]<-"Post"

#Create a New Data Set with an Average Depression Variable and Race
Mothership_Mean_RDQ<-Mothership_Long_RDQ %>%
  group_by(rm,PrePost) %>%
  summarise(RDQ=mean(RDQ,na.rm=TRUE))

#Add a Paired Variable to the New Data Set
Mothership_Mean_RDQ<-mutate(Mothership_Mean_RDQ,
                            paired=case_when(rm=="rdqPRE_coping"~"1",
                                             rm=="rdqPOST_coping"~"1",
                                             rm=="rdqPRE_pmh"~"2",
                                             rm=="rdqPOST_pmh"~"2",
                                             rm=="rdqPRE_func"~"3",
                                             rm=="rdqPOST_func"~"3",
                                             rm=="rdqPRE_wbs"~"4",
                                             rm=="rdqPOST_wbs"~"4",
                                             rm=="rdqPRE_sym"~"5",
                                             rm=="rdqPOST_sym"~"5"))

#Pre Post RDQ Plot
ggplot(data=Mothership_Mean_RDQ,
       aes(x=factor(PrePost,level=c("Pre","Post")),y=RDQ,
           group=rm,color=paired,!is.na(PrePost)))+
  geom_point(size=2.5)+
  geom_line(group=Mothership_Mean_RDQ$paired,size=1.5)+
  scale_x_discrete(name="Time Point")+
  scale_y_continuous(name="RDQ Subscore")+
  ggtitle("Mean Pre and Post Scores of the Remission from Depression Questionnaire")+
  theme(panel.grid.major.y = element_line(color = "grey",size = 0.5,linetype = 2))+
  theme(panel.background=NULL)+
  theme(legend.title = element_blank())+
  scale_color_manual(values=c("1"="#8956BB","2"="#4C4452","3"="#B1A8B9",
                              "4"="#00BCA3","5"="#008570"),
                     labels=c("Coping Skills","Positive Mental Health",
                              "Functioning","Well-Being","Symptoms"))

```

```{r, include=FALSE}
# Depression symptoms df
# CUDOS_Score <- select(Mothership,
#                       Days_complete_1,Day1_CUDOS:Day40_CUDOS)
# 
# # Anxiety symptoms df
# CUXOS_Score <- select(Mothership,
#                       Days_complete_1,Day1_CUXOS:Day40_CUXOS)

```

```{r, include=FALSE}
#1. First adjust the their values for length of stay above 40 to 40, because the measures don't go above 40 days.
# 
# 
# Mothership_post <- Mothership %>%
#   mutate(Days_complete_W = ifelse(Days_complete_1 > 40, 40,Days_complete_1))
# 
# 
# 
# Mothership <- Mothership %>%
#   mutate(Days_complete_W = ifelse(Days_complete_1 > 40, 40,Days_complete_1))
#          
# 
# Mothership$Day1_CUDOS
# Mothership$Days_complete_1
# Mothership$Days_complete_W
# 
# view(Mothership)
# #2. Then create df with CUDOS and CUXOS scores
#   # Depression
# 
#     CUDOS_Score <- select(Mothership,
#                       Day1_CUDOS:Day40_CUDOS)
# 
#   # Anxiety
#     CUXOS_Score <- select(Mothership,
#                       Day1_CUXOS:Day40_CUXOS)
# 
# 
# #3. Extract the valeues for each patient at their last day (length of stay)
#     ##### Why is it not includeing NAs.
#     
# 
# CUDOS_post = CUDOS_Score[cbind(seq_along(Mothership$Days_complete_W),Mothership$Days_complete_W)]
# CUXOS_post = CUXOS_Score[cbind(1:length(Mothership$Days_complete_W),Mothership$Days_complete_W)]
# 
# seq_along(Mothership$Days_complete_W)
# Mothership$Days_complete_W
# CUDOS_post
# 
# Mothership %>%
#   mutate(Dep_post = CUDOS_post)
# 
# 
# Mothership %>%
#   mutate(Dep_post = CUDOS_post)
# 
# CUDOS_post
# #4. Extract the value for their first day
# CUDOS_pre<-Mothership$Day1_CUDOS
# CUXOS_pre<-Mothership$Day1_CUXOS
# AnxDep_PrePost_test <- data.frame(cbind(Mothership$ID1,CUDOS_pre,CUDOS_post,CUXOS_pre,CUXOS_post))
# AnxDep_PrePost_test
# colnames(AnxDep_PrePost)[1] = "ID1"
# 
# AnxDep_PrePost_test
# ## This is where I am Figure out where to go from here *******************
# #5. Create long dataframe
# 
# AnxDep_PrePost_Lon <- gather(AnxDep_PrePost,ID1:CUXOS_post,factor_key = "T")
# 
# AnxDep_PrePost_Long <- pivot_longer(AnxDep_PrePost,ID1:CUXOS_post)
# 
# 
# AnxDep_PrePost_Long
# ?pivot_longer()
# 
# AnxDep_PrePost_Long <- AnxDep_PrePost_Long[order(AnxDep_PrePost_Long$ID1),]
# 
# Mothership_Long_RDQ <- gather(Mothership_RDQ,rm,RDQ,rdqPRE_coping:rdqPOST_sym,
#                               factor_key = "T")
# Mothership_Long_RDQ

```

```{r, include=FALSE}
# Mothership_Post <- Mothership %>%
#   filter(!is.na(Days_complete_1)) %>%
#   mutate(Days_complete_W = ifelse(Days_complete_1 > 40, 40,Days_complete_1)) 
#   
# 
# Mothership_Post$Days_complete_W
# 1:length(Mothership_Post$Days_complete_W)
# 
# # Depression
# CUDOS_Score <- select(Mothership_Post, Day1_CUDOS:Day40_CUDOS)
# 
# CUDOS_Score
# 
# 1:length(Mothership_Post$Days_complete_W)
# 
# # Anxiety
# CUXOS_Score <- select(Mothership_Post, Day1_CUXOS:Day40_CUXOS)
# 
# values <- cbind(seq_along(Mothership_Post$Days_complete_W),Mothership_Post$Days_complete_W[1:8610])
# values
# CUDOS_post = CUDOS_Score[cbind(seq_along(Mothership_Post$Days_complete_W)[1:8610],Mothership_Post$Days_complete_W[1:8610])]
# CUDOS_post
# 
# 
# CUDOS_post = CUDOS_Score[cbind(1:length(Mothership_Post$Days_complete_W),Mothership_Post$Days_complete_W)]
# CUXOS_post = CUXOS_Score[cbind(1:length(Mothership_Post$Days_complete_W),Mothership_Post$Days_complete_W)]
# 
# CUDOS_post
# 
# values <- cbind(1:length(Mothership_Post$Days_complete_W),Mothership_Post$Days_complete_W)
# values
# Mothership_Post<- Mothership_Post %>%
#   mutate(Dep_post = CUDOS_post)
```